R Script V5

### 2023-05-12 Data & Scripts/MSC R Script.Rmd
March 12th
New version of the R script + PCAs. Includes pathway analyses.

---
title: "R Script V4"
author: "Alex Popescu"
date: "2023-01-23"
output: github_document
editor_options: 
  chunk_output_type: console
---

V4 - Made some revisions and added plots. All colors are colorblind friendly.

@@

> The number of crows in the area is not a good estimate of group size. The number of foraging members during the foraging event should be better.
> Whether or not the site was baited will likely affect the duration of movement, since crows will need to look for food in non-baited recordings, while the crows have easy access to a large concentration of food in baited recordings. Therefore, the baited crows are expected to move less than non-baited crows.
> #KMG: I can imagine a reviewer would want to know hwo the baiting was standardized - something to keep in mind when you are writing. I removed the observations in agricultural settings because they had no sentinels present. This could be talked about in the discussion: agricultural fields may lack perches for individuals to act as sentinels. Vineyards may have some structures, but the visibility is limited when the vines are laden with fruit.\
> #KMG: This is something that can go in the methods. You had three sites, but one was removed due to lack of sentinals, and then you can have a bit in the discussion. I removed observations past Sept 29 because I was told behaviors can change wildly in the winter. The number of crows spotted decreased past that date, suggesting the migrants have left or were leaving. To not have to include time of year (I already have part of the breeding season), I decided to simplify.\

@@

```{r Packages, include=F}
library("tidyverse")
library("psych")
library("car")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
```

@@

```{r Open Crow and manage, include =F}
setwd(getwd())
Crow<-read.csv("Crow.csv", stringsAsFactors = T)
#Some management #
Crow$ID<-as.factor(Crow$ID)
Crow$Number.of.Crows<-as.numeric(Crow$Number.of.Crows)
Crow$LDur<-log(Crow$Duration)
```

@@

> I first decided to look at the distribution of my durations. I set up a histogram with the distibution of durations across all behaviors. I had to log-transform the duration to get a better distribution.\

@@

```{r Crow boxplot, echo = F}
Box<-ggplot(data = Crow)+
  geom_boxplot_pattern(aes(x = G.Env, y = LDur, fill = Behavior, pattern = Sentinel)
                       , pattern_angle = 45
                       , pattern_size = 0.05
                       , pattern_spacing = 0.02
                       , color = "black")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  scale_pattern_manual(values = c("crosshatch", "none")
                       , labels = c("Sentinel Absent", "Sentinel Present"))+
  guides(pattern = guide_legend(title = "Presence of a sentinel"
                                , override.aes = list(fill = c("white")))
         , fill = guide_legend(override.aes = list(pattern = "none")))+
  xlab("Generalized Environment")+
  ylab("Log Duration(s)") +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())
Box
#KMG suggestion: I'd make the hatch marks the one where sentinal is absent. They remind me of X's which makes me think there is no sentinal present. Just me, though, so not a big deal. Also a reminder to check if these are colour blind firendly-I don't think the blue and green are.

#AP: I swapped the crosshatch in the figure and used a colorblind-friendly palette.
```

@@

> #KMG: Is there a way to group the violin plot like the boxplot? So the violins for foraging sentinal present/absent are next to each other? Will be easier to see your conclusions. Also, have you looked at plotting means and SE? Here's a plot #Kiyoko plot of means and SE. KMG: I've put the function to generate means, SE, SD, and CI in the github. With this plot, it's easier to see the differences in the commercial area where the sentinal presence increases the behaviours. It's much messier in the green areas. I wonder if it's a refuge or visibility thing?

```{r Base R Plot, include = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
MeanDur <- summarySE(data = Crow, measurevar = "Duration"
                     , groupvars = c("G.Env", "Behavior", "Sentinel"))
#switching the order of groupvars will change how they are grouped.
plot(x = c(1:12)
     , y = MeanDur$Duration
     , xlim = c(0,12)
     , ylim = c(1,3)
     , type = "n"
     , ylab = "Duration (seconds)", xlab = "Generalized Environment")
arrows(c(1:12), MeanDur$Duration - MeanDur$se
       , c(1:12), MeanDur$Duration + MeanDur$se
       , length = 0.1, angle = 90, code = 3, lwd = 2)
points(x = c(1:12), y = MeanDur$Duration
       , pch = c(rep(c(16, 17), 3)), cex = 2
       , col = c("orange", "orange", "blue", "blue", "green", "green"))
legend(x = "bottomright", c("Sentinel Absent", "Sentinal Present", "Foraging", "Alert", "Moving")
       , bty = "o", cex = 1.05, bg = "white"
       , pch = c(16, 17, 15, 15, 15), col = c("black", "black", "orange", "blue", "green"))
#KMG: I'm not going to fix the X-axis labels and those kinds of things right now. I can do that if you like this plot or things it's informative. I think it's easier to see the differences. left side is commercial, right side is green. Legend should be correct.
```

@@

> Very cool! I modified the violin and boxplot, but I think the last plot is the most informative. We can see that the presence of a sentinel increases the duration of bouts in commercial areas. In green areas, however, this effect is reduced apart from in alert behavior.

@@

@@

> I used robustlmm::rlmer to create the models. It takes longer to run than the regular lmer test because of the computations it makes to weigh the model. I then plotted the models to test the assumptions of the model. The model I used was "log(Duration) \~ Behavior + (Sentinel\* G.Env) + Number of crows + Baited + (1\|ID)". KMG: what is the backslash?
>
> Behavior is expected to be affected by both the presence of a sentinel and the environment the crows are in. I am therefore curious to see the interactions between these factors. The presence of bait and number of crows are not linked with the presence of a sentinel or the environment. Arguably, larger groups are more likely to have sentinels than not, but I don't believe the interaction to be important here.

@@

```{r Duration RLMM, echo=F}
RLMM<-rlmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
plot(RLMM)
```

> That took AGES to run. There's a big "cliff" on the normal Q-Q vs residuals plot. This means that there are more extreme values than would be expected if they came from a truly normal population.

### Bout Results

```{r Duration RLMM Summary, echo = F}
summary(RLMM)
```

@@

```{r Duration Better Summary, echo = F}
sjPlot::tab_model(RLMM
                  , show.re.var = T
                  , title = "Effects of generalized environment, bait, number of crows and the presence of a sentinel on durations of behaviors"
                  , dv.labels = " Effects of on duration of behaviors")
#KMG: is baiting not in here as well?
```

> Like this! This is much better! Lets try a simpler output. Curious... Removing the three-way interaction caused many factors to become significant. Number of crows remains non-significant. #KMG: Can we visualize the the baited effects on duration as well?

```{r Simpler output, echo = F}
Simple.Dur<-lmer(LDur~Behavior+Sentinel*G.Env+Number.of.Crows + Baited + (1|ID), data = Crow)
Anova(Simple.Dur)
ranova(Simple.Dur)
```

> Simple model shows significant effects of Behavior, and the interaction between generalized environment and Behavior, and generalized environment and the presence of a sentinel.

@@

> I generated a plot to better see what the interaction between Generalized Environment and the presence of a sentinel. First, let's make a summary of the results.

```{r Duration Summary, echo=F}
MeanDur
```

> There are too many combinations to effectively see any trends, let's look at the significant interactions from the models.

```{r Bait effect on bout, echo = F}
BaitPlot<- Crow %>%
  summarySE(measurevar = "Duration"
            , groupvars = "Baited") %>%
  ggplot(aes(x = Baited, y = Duration))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  xlab("")+
  ylab("Mean Bout Duration (s)")+
  scale_x_discrete(labels=c("Not Baited", "Baited"))+
  theme(legend.position = 'bottom'
        , axis.title.x = element_blank())
BaitPlot
```

### Plot the Interactions

<!-- --- -->

<!-- ```{r Behavior x G.Env Plot, echo = F} -->

<!-- BGInteraction<-Crow %>% -->

<!--   group_by(Behavior, G.Env) %>% -->

<!--   summarize(mean = mean(log(Duration)) -->

<!--             , se = sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , upper = mean(log(Duration))+sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , lower = mean(log(Duration))-sd(log(Duration))/sqrt(length(log(Duration))) -->

<!--             , .groups = 'keep') %>% -->

<!--   ggplot(aes(x = G.Env, y = mean, colour = Behavior))+ -->

<!--   geom_point(position = position_dodge(width = 0.9)) + -->

<!--   geom_text(aes(label = signif(mean)) -->

<!--             , size = 3 -->

<!--             , position=position_dodge2(width=0.9, preserve='single') -->

<!--             , hjust = 1.1 -->

<!--             , vjust = -.5)+ -->

<!--   geom_errorbar(aes(ymin=lower, ymax=upper) -->

<!--                 , width = 0.2 -->

<!--                 , position = position_dodge(width=.9))+ -->

<!--   theme_bw()+ -->

<!--   theme(text=element_text(size = 14))+ -->

<!--   labs(colour = "Behavior")+ -->

<!--   xlab("")+ -->

<!--   ylab("Mean log Bout Duration")+ -->

<!--   scale_colour_manual(values = c("orange", "blue", "green") -->

<!--                     , labels = c("Foraging", "Alert", "Moving"))+ -->

<!--   scale_x_discrete(labels=c("Commercial", "Green Area"))+ -->

<!--   theme(legend.position = 'bottom') -->

<!-- BGInteraction -->

<!-- ``` -->

<!-- > The error bars here represent the standard error. It looks like in both environments, the duration of bouts of alertness are smaller than the other two behaviors. Interestingly, in Green areas, individuals decrease the durations of bouts of movement and increase the duration of bouts of foraging. Let's next check out the interaction between the presence of a sentinel and generalized environment. -->

```{r Sentinel x G.Env Plot, echo = F}
SGInteraction<- Crow %>%
  summarySE(measurevar = "Duration"
            , groupvars = c("G.Env", "Sentinel")) %>%
  ggplot(aes(x = G.Env, y = Duration, colour = Sentinel))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  labs(colour = "Sentinel")+
  xlab("")+
  ylab("Mean Bout Duration (s)")+
  scale_colour_manual(values = cbPalette
                    , labels = c("Sentinel Absent", "Sentinel Present"))+
  scale_x_discrete(labels=c("Commercial", "Green Area"))+
  theme(legend.position = 'bottom'
        , axis.title.x = element_blank())
SGInteraction
```

> Very cool! I believe I'm getting better at this (or I make good annotations on other scripts...). There is a difference in bout duration between foragers in the presence and absence of a sentinel. The difference is reduced in green areas. Again, the error bars are the standard error. KMG: Cool. You'll want to think about what exactly do you want to present. You can generate all sorts of figures. I personally think these means and SE plots are better. You can add the raw data in as a cloud of points, or you can add histograms sideways next to the plots as well.

### Effect Size

> I next plotted the effect size for each effect.

```{r Duration Effect Size, echo = F}
sjPlot::plot_model(RLMM
                   , show.values=T
                   , show.p=T
                   , value.offset = 0.25
                   , value.size = 4
                   , title = "Effects of generalized environment, number of crows and the presence of a sentinel on the duration of behaviors")
```

> Interesting... According to my model, Alert behavior had a significant negative effect on duration, while the presence of a sentinel and Green Areas have a significant positive effect on duration. The interaction between Moving and Green areas also had a significant negative effect on bout duration. I'm still relatively green when it comes to LMMs, so this may be evident, but I'll make my assumptions clear. From the previous figure, I see that "Head Down", G.Env[Commercial], Baited[N], etc are missing. Therefore I assume that the comparisons are made with those. #KMG: THat is correct. Therefore, I can conclude that: - Crows spend significantly less time with their head's up than down. - Crows spend significantly less time moving than with their heads down. - The duration of behaviors decreases significantly in the absence of a sentinel. - The duration of behaviors decreases significantly in Commercial areas. - The duration of behaviors decreases significantly in baited sites. KMG: When you present these kinds of things, its helpful to write in the same direction. So instead of significantly less and increases, perhaps significantly more and increases, or significantly less and decreases. - There was an interaction between the presence of a sentinel and the generalized environment. That being said, many papers start removing variables to fine tune their model. This is something I am considering in order to better identify the effects of the presence of a sentinel and the type of environment in which the individual forages in.

### What about the group size?

> From the previous results, the number of crows had little effect on the individual vigilance of foragers. This is slightly unexpected, since we would assume there is safety in numbers. By that logic, and hypotheses like the 'many eyes' hypothesis, we would expect to observe a decrease in the duration of some behaviors, namely "Alert". To test this, I decided to plot the model's estimates on the duration of behavior "Alert" to see if there is any trend.

```{r Effects of Group Size, echo = F}
##### Plot estimates with data - Number of crows #####
effects_Num<-effects::effect(term="Number.of.Crows", mod = RLMM)
summary(effects_Num)
Num<-as.data.frame(effects_Num)
Num.crows<-ggplot() +
  geom_point(data = subset(Crow, Behavior == "HU")
             , aes(x = Number.of.Crows, y = log(Duration))
                   , color = "black") + 
  geom_point(data = Num
             , aes(x = Number.of.Crows, y = fit)
             , color = "blue") +
  geom_line(data = Num
            , aes( x = Number.of.Crows, y = fit, group = 1)
            , color = "blue") +
  geom_ribbon(data = Num
              , aes( x = Number.of.Crows, ymin = lower, ymax = upper, group = 1)
              , alpha = 0.3
              , fill = "blue") + 
  labs(x="Number of crows (centered and scaled)"
       , y = "log Duration (s)")
Num.crows
```

> Looks like the data does not follow the predictions from the model. I therefore conclude that the group size had no effect on the duration of alert behavior, at any group size (that I tested).

@@

```{r Proportion setup, include = F}
Proportion<-read.csv("Proportions.csv", stringsAsFactors = T)
Proportion$ID<-as.factor(Proportion$ID)
Proportion$AProp<-asin(sqrt(Proportion$Proportion))
```

```{r Proportion Double Check, echo = F}
str(Proportion)
```

@@

```{r Proportion Summary, echo = F}
MeanProp<-summarySE(Proportion, "Proportion", groupvars = c("G.Env", "Behavior", "Sentinel"))
MeanProp
```

@@

```{r Proportion Histograms Individual, include = F}
Untransformed.Prop<-hist(Proportion$Proportion)
Transformed.Prop<-hist(Proportion$AProp, main="Transformed")
```

@@

>>>>

---
title: "R Script V5"
author: "Alex Popescu"
date: "2023-05-12"
output: github_document
editor_options: 
  chunk_output_type: console
---

V5 - Major overhaul and data/analysis improvement.

@@

> The number of crows in the area is not a good estimate of group size. The number of foraging members during the foraging event should be better.
> Whether or not the site was baited will likely affect the duration of movement, since crows will need to look for food in non-baited recordings, while the crows have easy access to a large concentration of food in baited recordings. Therefore, the baited crows are expected to move less than non-baited crows.
>I removed the observations in agricultural settings because they had no sentinels present. This could be talked about in the discussion: agricultural fields may lack perches for individuals to act as sentinels. Vineyards may have some structures, but the visibility is limited when the vines are laden with fruit.\
>I removed observations past Sept 29 because I was told behaviors can change wildly in the winter. The number of crows spotted decreased past that date, suggesting the migrants have left or were leaving. To not have to include time of year (I already have part of the breeding season), I decided to simplify.\

@@

```{r Packages, include=F}
library("tidyverse")
library("psych")
library("car")
library("ggplot2")
library("ggpattern")
library("lmerTest")
library("robustlmm")
library("sjPlot")
library("effects")
library("glmm")
library("doBy")
library("FactoMineR")
library("factoextra")
```

@@

```{r Open Crow and manage, include =F}
setwd(getwd())
Crow<-read.csv("Crow.csv", stringsAsFactors = T)
#Some management #
Crow$ID<-as.factor(Crow$ID)
Crow$Number.of.Crows<-as.numeric(Crow$Number.of.Crows)
Crow$CODED_BAIT<-as.numeric(Crow$CODED_BAIT)
Crow$CODED_G.ENV<-as.numeric(Crow$CODED_G.ENV)
Crow$CODED_NB_CROWS<-as.numeric(Crow$CODED_NB_CROWS)
Crow$CODED_SENTINEL<-as.numeric(Crow$CODED_SENTINEL)
Crow$CODED_BEHAVIOR<-as.numeric(Crow$CODED_BEHAVIOR)
Crow$LDur<-log(Crow$Duration+1)
```

@@

> I first decided to look at the distribution of my durations. I set up a histogram with the distibution of durations across all behaviors. I had to log-transform the duration to get a better distribution (log(Duration+1)).\

@@

```{r Crow boxplot, echo = F}
Box<-ggplot(data = Crow)+
  geom_boxplot_pattern(aes(x = G.Env, y = LDur, fill = Behavior, pattern = Sentinel)
                       , pattern_angle = 45
                       , pattern_size = 0.05
                       , pattern_spacing = 0.02
                       , color = "black")+
  theme(legend.position = "bottom")+
  scale_fill_manual(values = cbPalette
                    , labels = c("Foraging", "Alert", "Moving"))+
  scale_pattern_manual(values = c("crosshatch", "none")
                       , labels = c("Sentinel Absent", "Sentinel Present"))+
  guides(pattern = guide_legend(title = "Presence of a sentinel"
                                , override.aes = list(fill = c("white")))
         , fill = guide_legend(override.aes = list(pattern = "none")))+
  xlab("Generalized Environment")+
  ylab("Log Duration(s)") +
  theme(axis.title.x = element_blank()
        , legend.position = "bottom"
        , legend.box="vertical"
        , legend.margin=margin())
Box
```

@@

>These plots are nice, but it is hard to really see the underlying trends. To best visualise them, I'll next plot a dot plot.

### The Dot Plot

```{r Base R Plot, include = F}
source("./Calculate Summary Table with SE, SD, CI by grouping variables.R")
MeanDur <- summarySE(data = Crow, measurevar = "Duration"
                     , groupvars = c("G.Env", "Behavior", "Sentinel"))

plot(x = c(1:12)
     , y = MeanDur$Duration
     , xlim = c(0,12)
     , ylim = c(1,3)
     , type = "n"
     , ylab = "Duration (seconds)", xlab = "Generalized Environment")
arrows(c(1:12), MeanDur$Duration - MeanDur$se
       , c(1:12), MeanDur$Duration + MeanDur$se
       , length = 0.1, angle = 90, code = 3, lwd = 2)
points(x = c(1:12), y = MeanDur$Duration
       , pch = c(rep(c(16, 17), 3)), cex = 2
       , col = c("orange", "orange", "blue", "blue", "green", "green"))
legend(x = "bottomright", c("Sentinel Absent", "Sentinal Present", "Foraging", "Alert", "Moving")
       , bty = "o", cex = 1.05, bg = "white"
       , pch = c(16, 17, 15, 15, 15), col = c("black", "black", "orange", "blue", "green"))
#KMG: I'm not going to fix the X-axis labels and those kinds of things right now. I can do that if you like this plot or things it's informative. I think it's easier to see the differences. left side is commercial, right side is green. Legend should be correct.
```

@@

> Very cool! I think the last plot is the most informative. We can see that the presence of a sentinel increases the duration of bouts in commercial areas. In green areas, however, this effect is not observed. In fact, the bouts of Alert behavior seem significantly reduced in green areas and in the presence of a sentinel.

@@

### PCAs
```{r Bout Duration PCA packages, include=F}
library("FactoMineR")
library("factoextra")
```
```{r Bout Duration PCA Book keeping, include=F}
PCA.Crow.Data<-Crow[,c(4,6,8,10,12:13)]
head(PCA.Crow.Data)
```
```{r Bout Duration PCA, echo=F}
PCA.Crow<-PCA(PCA.Crow.Data[,-5], ncp=5)
```

>I'm new to PCAs, so bear with me please. From what I see, presence of bait and sentinel, and the number of crows all contibute to PC1, while Generalized environment seems to be contributing to PC2. The behavior appears to contribute to PC2, but not a lot.

#### Eigenvalues

>Let's have a look at the eigenvalues.

``` {r Bout Duration PCA eigenvalues, echo = F}
Crow.eig.val<-get_eigenvalue(PCA.Crow)
Crow.eig.val
```

>Pretty small values. The second column shows how much contribution each dimension has to explain the total variance in the data. Let's quickly plot that.

#### Scree Plot

```{r Bout Duration PCA Scree Plot, echo = F}
fviz_eig(PCA.Crow, addlabels=T, ylim = c(0,40))
```

>Pretty cool. The first 3 PCs explain ~77% of the variance. Adding the 4th PC explains ~93% of the variation.

#### Contribution of variables to Principal Components

```{r Bout Duration PCA var results, include = F}
Crow.var<-get_pca_var(PCA.Crow)
```
```{r Bout Duration PCA Contribution Dim 1, echo=F}
fviz_contrib(PCA.Crow, choice="var", axes = 1, top = 20)
```

>The red dotted line represents the expected average contribution, assuming the contribution of the variables were uniform. In other words, the line is equal to 1/the number of variables; in this case 1/5 or 20%.

> From this plot, I can see that, as I saw previously in the correlation circle, the number of crows, the presence of a sentinel and the presence of bait contribute the most to PC1 (Dimension 1)

```{r Bout Duration PCA Contribution Dim 2, echo=F}
fviz_contrib(PCA.Crow, choice="var", axes = 2, top = 20)
```

> The generalized environment contributes the most to PC2.

```{r Bout Duration PCA Contribution Dim 3, echo=F}
fviz_contrib(PCA.Crow, choice="var", axes = 3, top = 20)
```

>The coded behavior contributes most to PC3.

```{r Bout Duration PCA Contribution Dim 4, echo=F}
fviz_contrib(PCA.Crow, choice="var", axes = 4, top = 20)
```

>OK! Now we're getting to the interesting bits. The presence of bait and the presence of a Sentinel contribute most to PC4. This makes me curious to see if there is an interaction between the two factors.

```{r Bout Duration PCA Contribution Dim 5, echo=F}
fviz_contrib(PCA.Crow, choice="var", axes = 5, top = 20)
```

>>The presence of a sentinel and the size of the group contribute most to PC 5. I believe this would imply that the interaction between the two could be interesting to look at. Considering these two variables are also major contributors to PC1, this makes the inclusion of the interaction in the model a must.

>From this, I believe I should look at the interaction between NB_Crows, Presence of a sentinel, and the presence of bait.

#### Plotting with Behaviors

>I'd like to plot the dimensions using the behaviors of the bouts.

```{r Bout Duration PCA Plot with Behavior PC1-2, echo=F}
fviz_pca_ind(PCA.Crow
             , geom.ind="point"
             , col.ind=PCA.Crow.Data$Behavior
             , palette = cbPalette
             , addEllipses = T
             , legend.title = "Behaviors"
             )
```

>Messy. The mean points of the behaviors are relatively centered.

```{r Bout Duration PCA Plot with Behavior PC2-3, echo=F}
fviz_pca_ind(PCA.Crow
             , axes = c(2,3)
             , geom.ind="point"
             , col.ind=PCA.Crow.Data$Behavior
             , palette = cbPalette
             , addEllipses = T
             , legend.title = "Behaviors"
             )
```

>Very clear separation by behavior type, with 'Head Up' shifted upward, 'Head Down' shifted downward, and 'Moving' in the middle.

```{r Bout Duration PCA Plot with Behavior PC1-3, echo=F}
fviz_pca_ind(PCA.Crow
             , axes = c(1,3)
             , geom.ind="point"
             , col.ind=PCA.Crow.Data$Behavior
             , palette = cbPalette
             , addEllipses = T
             , legend.title = "Behaviors"
             )
```

>Good separation, but the concentration ellipses have increased in size. Curious.

#### Biplot

>Finally, let's look at the biplot. From what I understand, it is essentially the correlation circle overlaid onto the data.

```{r Bout Duration PCA Biplot PC1-2, echo=F}
fviz_pca_biplot(PCA.Crow
                ,geom.ind = "point"
                , repel = T
                , col.ind=PCA.Crow.Data$Behavior
                , addEllipses = T
                , label = "var"
                , palette=cbPalette
                , col.var="black"
                , legend.title = "Behavior")
```
```{r Bout Duration PCA Biplot PC2-3, echo=F}
fviz_pca_biplot(PCA.Crow
                , axes = c(2,3)
                , geom.ind = "point"
                , repel = T
                , col.ind=PCA.Crow.Data$Behavior
                , addEllipses = T
                , label = "var"
                , palette=cbPalette
                , col.var="black"
                , legend.title = "Behavior")
```
```{r Bout Duration PCA Biplot PC1-3, echo=F}
fviz_pca_biplot(PCA.Crow
                , axes = c(1,3)
                , geom.ind = "point"
                , repel = T
                , col.ind=PCA.Crow.Data$Behavior
                , addEllipses = T
                , label = "var"
                , palette=cbPalette
                , col.var="black"
                , legend.title = "Behavior")
```

>Nice! I think I've done the PCA analysis to the best of my abilities, but I am very much open to comments and feedback!

>From what I can gather, I should model the data as follows:/
>LDur~Behavior+NB_Crows*Bait*Sentinel+G.Env+(1|ID)

@@

> I used robustlmm::rlmer to create the models. It takes longer to run than the regular lmer test because of the computations it makes to weigh the model. I then plotted the models to test the assumptions of the model.

> Behavior is expected to be affected by both the presence of a sentinel and the environment the crows are in. I am therefore curious to see the interactions between these factors. The presence of bait and number of crows are not linked with the presence of a sentinel or the environment. Arguably, larger groups are more likely to have sentinels than not, but I don't believe the interaction to be important here.

@@

```{r Bout Duration Model Book keeping, include=F}
Crow$CODED_NB_CROWS<-as.factor(Crow$CODED_NB_CROWS)
```

```{r Duration RLMM, echo=F}
RLMM.Dur.I<-rlmer(LDur~Behavior+Sentinel*CODED_NB_CROWS*Baited+G.Env+ (1|ID), data = Crow)
plot(RLMM.Dur.I)
```

> That took pretty long to run. Let's next see the results of the model. I don't see huge deviations from the assumptions. According to the error message, my fixed-effect model matrix is rank deficient. I don't believe I need to worry too much about this warning.

```{r Duration RLMM Summary, echo = F}
summary(RLMM.Dur.I)
```

@@

```{r Duration Better Summary, echo = F}
sjPlot::tab_model(RLMM.Dur.I
                  , show.re.var = T
                  , title = "Model Results with Interactions"
                  , dv.labels = " Effects of on duration of behaviors")

>Oh fun! Adding the three-way interaction caused all but the behavior to be non-significant. Let's try removing the interactions.

```{r Duration RLMM No Interactions, echo=F}
RLMM.Dur.NoI<-rlmer(LDur~Behavior+Sentinel+CODED_NB_CROWS+Baited+G.Env+ (1|ID), data = Crow)
sjPlot::tab_model(RLMM.Dur.NoI
                  , show.re.var = T
                  , title = "Model Results without Interactions"
                  , dv.labels = " Effects of on duration of behaviors")
```

>Same result!

>Let's try the good old regular lmer and see if that changes anything.

```{r Simpler output, echo = F}
Simple.Dur.I<-lmer(LDur~Behavior+Sentinel*CODED_NB_CROWS*Baited + G.Env + (1|ID), data = Crow)
Simple.Dur.NoI<-lmer(LDur~Behavior+Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data = Crow)
anova(Simple.Dur.NoI, Simple.Dur.I)
```

>Using the simple model, I looked at the AIC values. The model without interactions had the lowest AIC value, therefore is preferred.Let's actually see what removing some factors in the model does to the AIC.

#### AIC Testing

```{r Bout Duration Drop1,echo=F}
drop1(Simple.Dur.NoI)
```

>Seems like dropping any of the variables will have no effect on the model's results. In other words, I can drop all factors except for 'behavior'.

```{r Bout Duration AIC testing}
Simplest.Dur<-lmer(LDur~Behavior+(1|ID), data=Crow)
anova(Simple.Dur.NoI, Simplest.Dur)
```

>Yep! The simplest model (LDur~Behavio+(1|ID)) is preferred. Let's try subsetting the data.

#### Subset Data
```{r Bout Duration Subset, include=F}
HU.Dur<-Crow[which(Crow$Behavior=="HU"),]
HD.Dur<-Crow[which(Crow$Behavior=="HD"),]
M.Dur<-Crow[which(Crow$Behavior=="M"),]
```
```{r Simple model on Subset HU data, echo=F}
Simple.HU.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =HU.Dur)
Anova(Simple.HU.Dur)
```

>Simple model on Head Up subset with no interactions shows no significance.

```{r Simple model on Subset HD data, echo=F}
Simple.HD.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =HD.Dur)
Anova(Simple.HD.Dur)
```

>Simple model on Head Down subset shows a significant effect of foraging group size, presence of bait and generalized environment.

```{r Simple model on Subset M data, echo=F}
Simple.M.Dur<-lmer(LDur~Sentinel+CODED_NB_CROWS+Baited + G.Env + (1|ID), data =M.Dur)
Anova(Simple.M.Dur)
```

> Simple model shows no significant effects.

@@

>Ok! Considering only the 'Head Down' subset had significant effects, let's plot those results.

#### Bait

```{r Bait effect on HD bout, echo = F}
BaitPlot<- HD.Dur %>%
  summarySE(measurevar = "Duration"
            , groupvars = "Baited") %>%
  ggplot(aes(x = Baited, y = Duration))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  xlab("Presence of Bait")+
  ylab("Mean Head Down Bout Duration (s)")+
  scale_x_discrete(labels=c("Not Baited", "Baited"))+
  theme(legend.position = 'bottom'
        )
BaitPlot
```

>Fun! The plot shows overlap between the standard error bars. In the presence of bait, crows will decrease the duration of their bouts of 'Head Down'.

#### G.Env

```{r G.Env effect on HD bout, echo = F}
EnvPlot<- HD.Dur %>%
  summarySE(measurevar = "Duration"
            , groupvars = "G.Env") %>%
  ggplot(aes(x = G.Env, y = Duration))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  xlab("Generalized Environment")+
  ylab("Mean Head Down Bout Duration (s)")+
  theme(legend.position = 'bottom'
        )
EnvPlot
```

>Very nice! We can see that in Green Areas, the bout duration of head down behavior is increased. The error bars, as before, are standard error.

#### Foraging Group Size

```{r NB_Crows effect on HD bout, echo = F}
NBPlot<- HD.Dur %>%
  summarySE(measurevar = "Duration"
            , groupvars = "CODED_NB_CROWS") %>%
  ggplot(aes(x = CODED_NB_CROWS, y = Duration))+
  geom_point(position = position_dodge(width = 0.9)
             , size = 3) +
  geom_text(aes(label = signif(Duration))
            , size = 3
            , position=position_dodge2(width=0.9, preserve='single')
            , hjust = -0.5
            , vjust = 0.1)+
  geom_errorbar(aes(ymin=(Duration-se), ymax=(Duration+se))
                , width = 0.2
                , position = position_dodge(width=.9))+
  theme_bw()+
  theme(text=element_text(size = 14))+
  xlab("Group Size")+
  ylab("Mean Head Down Bout Duration (s)")+
  scale_x_discrete(labels=c("Small", "Large"))+
  theme(legend.position = 'bottom'
        )
NBPlot
```

>Very nice! In larger foraging groups, individuals increase their bouts of head down behavior.

>I can therefore make the following conclusions:
>-Crows do not modify the duration of bouts of head up or moving
>-Crows will modify the duration of bouts of head down behavior.
>-In larger groups, crows will increase the duration of bouts of 'Head Down", possibly due to the many eyes hypothesis.
>-In the presence of Bait, crows will decrease the duration of bouts of 'Head Down'. This could be due to the larger concentration of food in the patch, and therefore less time required to look for food. 
>-In Green Areas, crows will increase the duration of bouts of 'Head Down'.

@@

```{r Proportion setup, include = F}
DATA<-read.csv("DATA.csv", stringsAsFactors = T)
DATA$ID<-as.factor(DATA$ID)
DATA$ASIN_PROPORTION<-asin(sqrt(DATA$BEHAVIOR_PROPORTION))
```

```{r Proportion Double Check, echo = F}
str(DATA)
```

@@

```{r Proportion Summary, echo = F}
MeanProp<-summarySE(DATA, "BEHAVIOR_PROPORTION", groupvars = c("GENERALIZED_ENVIRONMENT", "BEHAVIOR", "SENTINEL_PRESENCE"))
MeanProp
```

@@

```{r Proportion Histograms Individual, include = F}
Untransformed.Prop<-hist(DATA$BEHAVIOR_PROPORTION, breaks=25)
Transformed.Prop<-hist(DATA$ASIN_PROPORTION, breaks=25)
```

@@

## PCAs

>Let's run another PCA

```{r DATA PCA SHEET, include=F}
PCA.DATA.SHEET<-DATA[,c(5,7,10,14,16,18,22,23,24,39)]
head(PCA.DATA.SHEET)
str(PCA.DATA.SHEET)
```
```{R DATA PCA, echo=F}
PCA.DATA<-PCA(PCA.DATA.SHEET[,c(-8,-9)])
```

>There are many variables here. I won't bother describing them here.

>Since the presence of a sentinel is not "God-Given", we'll run another PCA without that factor.

```{r DATA PCA NO SENTINEL, echo=F}
PCA.DATA.NOSENT<-PCA(PCA.DATA.SHEET[,c(-5,-8,-9)])
```

>Very similar to the one above.

### Scree Plot

```{r DATA PCA Scree Plot, echo = F}
get_eigenvalue(PCA.DATA)
fviz_eig(PCA.DATA, addlabels=T, ylim = c(0,40))
```

>The first PC covers ~40% of the variation. The first 4 PCs cover ~82% of the variation.
